Outline

### 2021-07-30 text/main.Rmd
text edits rev minor

**Abstract:**
Preserving the functioning of coral reefs is a critical challenge of the 21st century. However, a lack of quantitative assessments of multiple functions across large spatial scales has hindered local and regional conservation efforts. We integrate empirically-parametrized bioenergetic models and global community surveys to quantify five key functions mediated by coral reef fishes. We show that functions exhibit critical trade-offs driven by diverging community structures, such that no reef can holistically maximize functioning. Further, functions are locally dominated by few species, but worldwide, the identity of dominant species greatly varies; 70% of the 1,110 species in our dataset are functionally dominant. Our results underline the need for a nuanced approach to coral reef conservation that considers multiple functions beyond the effect of standing stock biomass.

@@

Coral reefs are among the most diverse and productive ecosystems on Earth and provide essential ecosystem services [@Teh2013].
Yet, their integrity is threatened by a plethora of anthropogenic stressors [@Hughes2017].
Severe declines in habitat quality and fish biomass have brought coral reef functioning to the forefront of scientific discourse [@Bellwood2019; @Brandl2019front].
However, our capacity to evaluate reef functioning typically relies on static proxies of functions, such as live coral cover, standing stock biomass of reef fishes, or the diversity of qualitative traits [e.g. @Cinner2020].
Conversely, we know comparatively little about actual functions - fluxes of elements and energy - and their drivers (but see [@Allgeier2016]), which constitutes a severe limitation to mefficient anagement [@Brandl2019front].

@@

The five key ecosystem functions performed by fishes across the world's reefs exhibit high variability (Fig. 1).
Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we observed a predictably strong relationship between fish biomass and all five functions (Fig. S1a-e, Fig. S2).
However, our analyses demonstrate striking variability after accounting for biomass: in communities with similar biomass, functions may differ by two orders of magnitude or more (Fig. S1a-f).
Thus, using biomass as a proxy for functioning masks fundamental differences in critical community-level functions.
Further, we demonstrate strong trade-offs among the five functions, independent of biomass (Fig.1, Fig. S1g).
For example, high herbivory rates and nitrogen excretion negatively correlate with rates of phosphorus excretion.
Consequently, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can holistically maximize functioning (Fig. 1).

>>>>

**Abstract:**
Preserving the functioning of coral reefs is a critical challenge of the 21st century. However, a lack of quantitative assessments of multiple functions across large spatial scales has hindered local and regional conservation efforts. We integrate empirically-parametrized bioenergetic models and global community surveys to quantify five key functions mediated by coral reef fishes. We show that functions exhibit critical trade-offs driven by varying community structures, such that no reef can holistically maximize functioning. Further, functions are locally dominated by few species, but worldwide, the identity of dominant species greatly varies; 70% of the 1,110 species in our dataset are functionally dominant. Our results underline the need for a nuanced approach to coral reef conservation that considers multiple functions beyond the effect of standing stock biomass.

@@

Coral reefs are among the most diverse and productive ecosystems on Earth and provide essential ecosystem services [@Teh2013].
Yet, their integrity is threatened by a plethora of anthropogenic stressors [@Hughes2017].
Severe declines in habitat quality and fish biomass have brought coral reef functioning to the forefront of scientific discourse [@Bellwood2019; @Brandl2019front].
However, our capacity to evaluate reef functioning typically relies on static proxies of functions, such as live coral cover, standing stock biomass of reef fishes, or the diversity of qualitative traits [e.g. @Cinner2020].
Conversely, we know comparatively little about actual functions - fluxes of elements and energy - and their drivers (but see [@Allgeier2016]), which constitutes a severe limitation to efficient management [@Brandl2019front].

@@

The five key ecosystem functions performed by fishes across the world's coral reefs exhibit high variability (Fig. 1).
Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we observed a predictably strong relationship between fish biomass and all five functions (Fig. S1a-e, Fig. S2).
However, our analyses demonstrate striking variability after accounting for biomass: in communities with similar biomass, functions may differ by two orders of magnitude or more (Fig. S1a-f).
Thus, using biomass as a proxy for functioning masks fundamental differences in critical community-level functions.
Further, we demonstrate strong trade-offs among the five functions, independent of biomass (Fig.1, Fig. S1g).
For example, high herbivory rates and nitrogen excretion negatively correlate with rates of phosphorus excretion.
Consequently, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can holistically maximize functioning (Fig. 1).
